\Title

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

The challenges are numerous in such situations and two major obstacles must be highlighted: the computational delay and the energy budget. First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun1998gradient}, then high definition images, such as VGG on Imagenet~\citep{Simonyan:2014cmh}. This has a cost and notably the fact that the computation time increases with the size of the image. This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards. However, the transition from images to videos, then to multiple cameras implies a further level in the change of scale in the amount of data processed. It also raises an issue that is often not addressed when it comes to image categorization. Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street. As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct response. This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes. %
Second, a major constraint for future systems is energy cost. As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption. In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures. As an example, the best performing model in the Imagenet challenge uses 2100 million parameters. Training the model typically consumes kilowatts of power and about 1 joule to infer a decision per single image. Less power-hungry solutions have emerged, such as Mobile Net~\citep{howard_searching_2019}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.

@@

\note{say that it is effortless in biology but difficult on conventional computers}
In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class. The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model. 
%
\subsubsection{Performance on synthetic data}
%
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.490\linewidth]{figures/figure_N_PGs.pdf}
    \includegraphics[width=0.490\linewidth]{figures/figure_N_PG_time.pdf}
	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
	    	     }
    \label{fig:model_results}
\end{figure}
% 
This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$. This assumes that we know the PGs as defined by the $W_b$ matrices. The underlying metric is the binary cross-entropy, as used in the logistic regression model. In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}. In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator. Using this forward model, it possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator. It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.

@@

%%%-----------------------------------------------------------------
%: fig:motion_task
%%%-----------------------------------------------------------------
\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motion_task.pdf}
    \note{TODO: redo figure according to caption}
    \caption{
    {\bf Motion detection task.} {\bf (a)}  {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye. To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory. {\bf (b)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation. Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events. This will constitute the input to the spiking neural network.}
    \label{fig:motion_task}
\end{figure}
%%%-----------------------------------------------------------------
Once these eye movement trajectories are generated, we can apply them to a visual scene. For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}. Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened). This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}. These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (see \seeFig{motion_task}-(b)). We will discretize the time in $1~\ms$ bins and produce movies of duration N-T = $250~\ms$. To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step. The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}. Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a motion of one pixel per frame (i.e., per time bin).

@@

%
Since the model is fully differentiable, we can now implement a semi-supervised learning rule. The loss function of the MLR model is the binary cross-entropy on the output of the classification layer and can be extended to space-time convolutions. Supervision was implemented using the input binary events as defined above and the labels as the desired output. The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions. As in this semi-supervised context, the label is lknown but the timing is not, we used a selection of the temporal suppport in an unsuperviesd manner. It is important to note that we weighted the cost function by considering only active cells, such that the error is only backpropagated to the spatial locations of these most active cells. This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier2007}. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We tested the effect of different parameters on a validation set to quantify the role of each parameter.

@@

Finally, the output of the MLR model is an event-based representation predicting at each position and at each time the probability of each motion. Such an output gives a form of optical flow that can be exploited for non-rigid motions, but we have defined here, for simplicity, an evaluation method that applies to our task with a full field motion. We have shown above that when different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities. By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output. This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error% $|| \speed - ||^2$ 
of the velocity estimate given by average estimate. 

@@

We first observed a high dependendance between the weights reaching the ON polarities and that reaching the OFF polarities. In particular, whenever a weight for a given position and delay is positive for one poalrity, it will be negative in the other. This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease. %ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to. 

@@

We plot here main axis of interests. First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy. This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}. \note{we should really do something more serious here...} Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
%
\subsection{Accuracy for the motion detection task}
%%%-----------------------------------------------------------------

Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection. We observed that the distribution of the kernels' weights is sparse, with most values near zero. As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses. Indeed, the computations are dominated by the convolution operation. In a dense setting, this corresponds for all voxels in the ouput to a sum over all voxels in the inputs for all weights in the kernel. If the support of information is sparse, then computations can be performed only on those events. Also, if we set some weights of the kernels to zero, then the sum can be skipped for thse adresses. Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.

@@

%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobilogical anatomical observations but also with the results from behavioral results. \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a spiking neural network model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.

>>>>

The challenges are numerous in such situations and two major obstacles must be highlighted: the computational delay and the energy budget. First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun1998gradient}, then high definition images, such as VGG on Imagenet~\citep{Simonyan:2014cmh}. This has a cost and notably the fact that the computation time increases with the size of the image. This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards. However, the transition from images to videos, then to multiple cameras implies a further level in the change of scale in the amount of data processed. It also raises an issue that is often not addressed when it comes to image categorization. Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street. As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct answer. This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes. %
Second, a major constraint for future systems is energy cost. As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption. In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures. As an example, the best performing model in the Imagenet challenge uses 2100 million parameters. Training the model typically consumes kilowatts of power and about 1 joule to infer a decision per single image. Less power-hungry solutions have emerged, such as Mobile Net~\citep{howard_searching_2019}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.

@@

\note{say that it is effortless in biology but difficult on conventional computers}
In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class. The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model. 
%
\subsubsection{Performance on synthetic data}
%
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.490\linewidth]{figures/figure_N_PGs.pdf}
    \includegraphics[width=0.490\linewidth]{figures/figure_N_PG_time.pdf}
	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
	    	     }
    \label{fig:model_results}
\end{figure}
% 
This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$. This assumes that we know the PGs as defined by the $W_b$ matrices. The underlying metric is the binary cross-entropy, as used in the logistic regression model. In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}. In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator. Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator. It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.

@@

%%%-----------------------------------------------------------------
%: fig:motion_task
%%%-----------------------------------------------------------------
\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motion_task.pdf}
    \note{TODO: redo figure according to caption}
    \caption{
    {\bf Motion detection task.} {\bf (a)}  {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye. To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory. {\bf (b)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation. Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events. This will constitute the input to the spiking neural network.}
    \label{fig:motion_task}
\end{figure}
%%%-----------------------------------------------------------------
Once these eye movement trajectories are generated, we can apply them to a visual scene. For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}. Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened). This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}. These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (\seeFig{motion_task}-(b)). We will discretize the time in $1~\ms$ bins and produce movies of duration N-T = $250~\ms$. To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step. The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}. Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a motion of one pixel per frame (i.e., per time bin).

@@

%
Since the model is fully differentiable, we can now implement a semi-supervised learning rule. The loss function of the MLR model is the binary cross-entropy on the output of the classification layer and can be extended to space-time convolutions. Supervision was implemented using the input binary events as defined above and the labels as the desired output. The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions. As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner. It is important to note that we weighted the cost function by considering only active cells, such that the error is only backpropagated to the spatial locations of these most active cells. This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier2007}. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We tested the effect of different parameters on a validation set to quantify the role of each parameter.

@@

Finally, the output of the MLR model is an event-based representation predicting at each position and at each time the probability of each motion. Such an output gives a form of optical flow that can be exploited for non-rigid motions, but we have defined here, for simplicity, an evaluation method that applies to our task with a full field motion. We have shown above that when different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities. By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output. This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error % $|| \speed - ||^2$ 
of the velocity estimate given by the average estimate. 

@@

We first observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities. In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative in the other. This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease. %ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to. 

@@

We plot here main axis of interests. First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy. This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}. \note{we should really do something more serious here...} Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
%
\subsection{Accuracy for the motion detection task}
%%%-----------------------------------------------------------------

Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection. We observed that the distribution of the kernels' weights is sparse, with most values near zero. As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses. Indeed, the computations are dominated by the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. If the support of information is sparse, then computations can be performed only on those events. Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses. Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.

@@

%
\section{Discussion}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results. \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations. 
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
To recap, we have introduced in this paper a spiking neural network model with heterogeneous delays that we have trained and evaluated on a complex motion detection task. The model was defined to optimally detect event-driven spatio-temporal motifs. We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example. We have then shown that the model displays similarities with the responses observed in biology. We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.
